---
title: "Document Title"
output: 
  html_document:
    code_download: yes
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE, message = FALSE, warning = TRUE)
```

Last updated: `r format(Sys.Date(), "%B %d %Y")`

See [Package References](https://dr-jt.github.io/semoutput/reference/index.html){target="_blank"} for help documentation on `semoutput` functions

**NOTE: We will use two data sets in this example output. They come from the `lavaan` tutorial data sets; the cfa data set is`HolzingerSwineford1939` and the sem data set is `PoliticalDemocracy`**

**NOTE: You can download the RMarkdown file associated with this output by selecting the 'Code' button on the top-right of this document and selecting "Download Rmd". You can also display the code used to produce each output by selecting the 'Code' buttons along the right of the document.**

**NOTE: For each CFA or SEM model output block there are 4 tabs:**

-   The "Summary Output" tab display nice looking tables summarizing the model results

-   The "Diagram Output" tab will display a model diagram

-   The "Residual Correlation Matrix" tab will display the residual correlation matrix

-   The "Full Output" tab will display the results from `summary()` along with parameter estimates and modification indices. This way you can still get the full output from a lavaan model as it provides more information than the "Summary Output". You can also add additional output to this section if you need more info about the model.

**Once you install the package, you will be able to access an Rmarkdown template by going to**:

File -\> New File -\> R Markdown... -\> From Template -\> CFA/SEM (lavaan)

# Setup

Required Packages

```{r warning=FALSE}
library(readr)
library(here)
library(dplyr)
library(lavaan)
library(psych)
library(semoutput)
library(semPlot)
library(sjPlot)
```

Import Data

```{r warning=FALSE}
## Import Data
# data <- read_csv(here("relative file path", "file name"))
cfa_data <- dplyr::select(HolzingerSwineford1939, -id, -school)
sem_data <- PoliticalDemocracy
```

```{r}
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Descriptives

**Typically there is only one descriptive table displayed but since we have two data sets for this one output two descriptive tables are displayed.**

```{r}
# Prints basic descriptive statistics
sem_descriptives(cfa_data)
sem_descriptives(sem_data)
```

------------------------------------------------------------------------

# Correlation Matrix

**There are also two correlation matrices**

**This is a publication quality correlation matrix that can be inserted into a manuscript.**

```{r}
# Uses sjPlot to print a nice looking correlation table
tab_corr(cfa_data, na.deletion = "pairwise", digits = 2, triangle = "lower")
tab_corr(sem_data, na.deletion = "pairwise", digits = 2, triangle = "lower")
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# EFA {.tabset .tabset-pills}

```{r}
efa_data <- dplyr::select(cfa_data, dplyr::starts_with("x"))

## Conduct EFA analysis with nfactors
efa_fit <- fa(efa_data, fm = "pa", nfactors = 3, rotate = "varimax")
```

## Summary Output

```{r}
efa_method(efa_fit)
efa_var(efa_fit)
efa_loadings(efa_fit)
efa_rotmatrix(efa_fit)
```

## Diagram Output

```{r}
fa.diagram(efa_fit)

## Determine the number of factors to extract
VSS.scree(efa_data)
fa.parallel(efa_data, fa = "fa")
VSS(efa_data, n = 4, rotate = "varimax")
```

## Full Output

```{r}
efa_fit
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# CFA {.tabset .tabset-pills}

```{r}
# specify the model
model <- '
# latent factors
visual =~ x1 + x2 + x3
textual =~ x4 + x5 + x6
speed =~ x7 + x8 + x9

# correlated errors

# constraints

'
# fit model
fit <- cfa(model = model, data = cfa_data, mimic = "lavaan", 
           estimator = "ML", missing = "ML", 
           std.lv = TRUE, std.ov = FALSE, test = "standard", 
           se = "standard", bootstrap = 1000)
```

## Summary Output

```{r}
sem_sig(fit)
sem_fitmeasures(fit)
sem_factorloadings(fit, standardized = TRUE, ci = "standardized")
sem_factorcor(fit)
```

## Diagram Output

```{r}
semPaths(fit, latents = factors, whatLabels = "std", layout = "tree2", 
         rotation = 2, style = "lisrel", optimizeLatRes = TRUE, 
         intercepts = FALSE, residuals = TRUE, curve = 1, curvature = 3, 
         sizeLat = 10, nCharNodes = 8, sizeMan = 11, sizeMan2 = 4, 
         edge.label.cex = 1.2, edge.color = "#000000")
```

## Residual Correlation Matrix

```{r}
sem_residuals(fit)
```

## Full Output

### Summary

```{r}
summary(fit, fit.measures = TRUE, standardized = TRUE)
```

### Parameter Estimates

```{r}
standardizedSolution(fit)
```

### Modification Indices

```{r}
modificationIndices(fit, sort. = TRUE, minimum.value = 3)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# SEM {.tabset .tabset-pills}

```{r}
# specify the model
model <- '
# measurement model
ind60 =~ x1 + x2 + x3
dem60 =~ y1 + y2 + y3 + y4
dem65 =~ y5 + y6 + y7 + y8

# regressions
dem60 ~ ind60
dem65 ~ ind60 + dem60

# covariances
y1 ~~ y5
y2 ~~ y4 + y6
y3 ~~ y7
y4 ~~ y8
y6 ~~ y8

# variances

'
# fit model
fit <- sem(model = model, data = sem_data, mimic = "lavaan", 
           estimator = "ML", missing = "ML", 
           std.lv = FALSE, std.ov = FALSE, test = "standard", 
           se = "standard", bootstrap = 1000)
```

## Summary Output

```{r}
sem_sig(fit)
sem_fitmeasures(fit)
sem_factorloadings(fit, standardized = TRUE, ci = "standardized")
sem_paths(fit, standardized = TRUE, ci = "standardized")
sem_factorcor(fit)
sem_factorvar(fit)
sem_rsquared(fit)
```

## Diagram Output

**Compared to the CFA figure above, I modified the r-code chunck option for figure width to `fig.width = 10` to make the image wider. And in the 'Code' I also changed the paramter for edge labels to `edge.label.cex = 8`.**

```{r fig.width=10}
semPaths(fit, latents = factors, whatLabels = "std", layout = "tree2", 
         rotation = 2, style = "lisrel", optimizeLatRes = TRUE, 
         intercepts = FALSE, residuals = TRUE, curve = 1, curvature = 3, 
         sizeLat = 10, nCharNodes = 8, sizeMan = 11, sizeMan2 = 4, 
         edge.label.cex = .8, edge.color = "#000000")
```

## Residual Correlation Matrix

```{r}
sem_residuals(fit)
```

## Full Output

### Summary

```{r}
summary(fit, fit.measures = TRUE, standardized = TRUE)
```

### Parameter Estimates

```{r}
standardizedSolution(fit)
```

### Modification Indices

```{r}
modificationIndices(fit, sort. = TRUE, minimum.value = 3)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Session Info

```{r comment=""}
citation("lavaan")
citation()
sessionInfo()
```
